### 2017-06-17 paper/lexicographic.tex
corrected some stuff about conditional expectation

\smallskip\noindent{\bf Conditional Expectation.} 
%The notion of conditional expectation 
%tightly connected to martingales. We present only a brief informal exposition 
%here, for formal treatment see, e.g.~\cite[Chapter 9]{Williams:book}.
Let $(\Omega,\mathcal{F},\probm)$ be a probability space, 
$X\colon\Omega\rightarrow 
\Rset$ an $\mathcal{F}$-measurable function, and $\mathcal{F}'\subseteq 
\mathcal{F}$ sub-sigma-algebra of $\mathcal{F}$. The \emph{conditional 
	expectation} of $X$ given $\mathcal{F}'$ is an $\mathcal{F}'$-measurable random 
variable denoted by $\E[X| \mathcal{F}']$ which satisfies, for each set $A\in 
\mathcal{F}'$, the following: 
\begin{equation}
\label{eq:cond-exp}
\E[X\cdot 1_A] = \E[\E[X|\mathcal{F}]\cdot 1_A],
\end{equation}
where $1_A \colon \Omega\rightarrow \{0,1\}$ is an \emph{indicator function} of 
$A$, i.e. function returning $1$ for 
each $\omega\in A$ and $0$ for each $\omega\in \Omega\setminus A$. Note that 
the left hand-side of~\eqref{eq:cond-exp} intuitively represents the expected 
value 
of $X(\omega)$ with domain restricted to $A$.

@@

\begin{definition}[Lexicographic Ranking Supermartingale]
Let $(\Omega,\genfilt,\probm)$ be a probability space, 
$\{\genfilt_i\}_{i=0}^{\infty}$ a filtration of $\genfilt$, $\stime$ a stopping 
time w.r.t. that filtration, and 
$\eps\geq 0$. 
An $n$-dimensional stochastic process $\{\vec{X}_{i}\}_{i=0}^{\infty}$ is a 
\emph{lexicographic $\eps$-ranking supermartingale for $\stime$ 
($\eps$-LexRSM)} if the 
following 
conditions hold:
\begin{compactenum}
\item For each $1\leq j \leq n$ the 1-dimensional stochastic process 
$\{\vecseq{X}{i}{j}\}_{i=0}^{\infty}$ is adapted to 
$\{\genfilt_i\}_{i=0}^{\infty}$.
%\item For each $i\in \Nset_0$ and each $1\leq j \leq n$ it holds 
%$\E[|\vecseq{X}{i}{j}|]<\infty$.
\item For each $\omega \in \Omega$ and $i\in \Nset_0$ it holds 
$\vec{X}_i (\omega)\geq \vec{0}$ (componentwise).
\item For each $i\in \Nset_0$ and each $\omega\in \Omega$ with 
$\stime(\omega)>i$ there exists $1\leq j 
\leq n$ such that
\begin{compactitem}
	\item $\E[\vecseq{X}{i+1}{j}\mid 
	\genfilt_i](\omega)\leq\vecseq{X}{i}{j}(\omega) - 
	\eps$; and
	\item for all $1 \leq j' < j$ it holds $\E[\vecseq{X}{i+1}{j'}\mid 
	\genfilt_i](\omega)\leq\vecseq{X}{i}{j'}(\omega)$.
\end{compactitem}
\end{compactenum}
\end{definition}

@@

For $\omega\in \Omega$ we define the level of $\omega$ at step $i$ to be the 
number $\levelrank{\omega}{i}$ equal to either $0$, if $\stime(\omega)\leq i$, 
or otherwise to the smallest $1\leq j \leq n$ such that 
$\E[\vecseq{X}{i+1}{j}\mid 
\genfilt_i](\omega)\leq\vecseq{X}{i}{j}(\omega)-\eps$. By the definition of 
$\eps$-LexRSM the value $\levelrank{\omega}{i}$ is well-defined for all 
$\omega$ and moreover, the random variable $\levelrank{}{i}$ is $\genfilt_i$-measurable. We denote by $\minlev(\omega)$ the smallest $0 \leq j \leq n$ such 
that $j$ is a level of $\omega$ at infinitely many steps. Note that $\omega \in 
\genRunSet_{\infty}$ if and only if $\minlev(\omega)\neq 0$. We denote by $M_i$ 
the set of all $\omega$'s with $\minlev(\omega)=i$.

>>>>

\smallskip\noindent{\bf Conditional Expectation.} 
%The notion of conditional expectation 
%tightly connected to martingales. We present only a brief informal exposition 
%here, for formal treatment see, e.g.~\cite[Chapter 9]{Williams:book}.
Let $(\Omega,\mathcal{F},\probm)$ be a probability space, 
$X\colon\Omega\rightarrow 
\Rset$ an $\mathcal{F}$-measurable function, and $\mathcal{F}'\subseteq 
\mathcal{F}$ sub-sigma-algebra of $\mathcal{F}$. A \emph{conditional 
	expectation} of $X$ given $\mathcal{F}'$ is an $\mathcal{F}'$-measurable random 
variable denoted by $\E[X| \mathcal{F}']$ which satisfies, for each set $A\in 
\mathcal{F}'$, the following: 
\begin{equation}
\label{eq:cond-exp}
\E[X\cdot 1_A] = \E[\E[X|\mathcal{F}]\cdot 1_A],
\end{equation}
where $1_A \colon \Omega\rightarrow \{0,1\}$ is an \emph{indicator function} of 
$A$, i.e. function returning $1$ for 
each $\omega\in A$ and $0$ for each $\omega\in \Omega\setminus A$. Note that 
the left hand-side of~\eqref{eq:cond-exp} intuitively represents the expected 
value 
of $X(\omega)$ with domain restricted to $A$.

Note that any $\mathcal{F}'$-measurable random variable satisfying~\eqref{eq:cond-exp} can be called a conditional expectation. The definition does not guarantee that the conditional expectation is uniquely defined or that it exists at all. However, from probability theory we have the following:

\begin{proposition}
\label{prop:conditional-exp-existence}
Let $(\Omega,\mathcal{F},\probm)$ be a probability space, 
$X\colon\Omega\rightarrow 
\Rset$ an $\mathcal{F}$-measurable function, and $\mathcal{F}'\subseteq 
\mathcal{F}$ sub-sigma-algebra of $\mathcal{F}$. Assume that one of the following conditions hold:
\begin{itemize}
\item $\E[|X|]<\infty$; or
\item $X$ is non-negative and for all $\omega\in\Omega$ it holds $X(\omega)\neq \infty$.
\end{itemize}
Then there exists a conditional expectation of $X$ given $\mathcal{F}'$ and it is almost-surely unique, i.e. for each two $\mathcal{F}'$-measurable functions $f$, $g$ that satisfy~\eqref{eq:cond-exp} it holds $\probm(\{\omega\mid f(\omega)\neq g(\omega)\})=0$.
\end{proposition}
\begin{proof}
The proof for the case when $\E[|X|]$ is standard and appears in many textbooks on probability theory (e.g.~\cite{Billingsley:book,Ash:book,Rosenthal:book}). The proof for the second case is essentially the same: the condition that $X$ is non-negative and not admitting infinite value suffices for satisfying the assumptions of Radon-Nikodym Theorem, the main theoretical tool used in the proof. For the sake of completeness we present the proof in \AppendixMaterial.
\end{proof}

Since the constraint~\eqref{eq:cond-exp} defining conditional expectation is phrased in terms of expected values, the almost-sure uniqueness cannot be strengthened to uniqueness, as re-defining a random variable on a set of probability zero does not change its expectation. In the following, when we say that a conditional expectation of a random variable $X$ satisfies some inequality (e.g. $\E[X\mid\mathcal{F}]\geq 0$) on set $L\subseteq \Omega$, we mean that for each $\mathcal{F}$-measurable function $\E[X\mid\mathcal{F}]$ satisfying~\eqref{eq:cond-exp} the inequality is satisfied for almost all $\omega \in L$.

@@

\begin{definition}
Let $(\Omega,\genfilt,\probm)$ be a probability space and $O\subseteq \Omega$ some set.
 We say that sets $L_1,\dots,L_n$ form an almost-sure partition of $O$ if $L_1,\dots,L_n$ are pairwise disjoint and $\probm(\{\omega\in O\mid \omega \not \in L_1\cup\dots \cup L_n\})=0$
\end{definition}

\begin{definition}[Lexicographic Ranking Supermartingale]
Let $(\Omega,\genfilt,\probm)$ be a probability space, 
$\{\genfilt_i\}_{i=0}^{\infty}$ a filtration of $\genfilt$, $\stime$ a stopping 
time w.r.t. that filtration, and 
$\eps\geq 0$. 
An $n$-dimensional stochastic process $\{\vec{X}_{i}\}_{i=0}^{\infty}$ is a 
\emph{lexicographic $\eps$-ranking supermartingale for $\stime$ 
($\eps$-LexRSM)} if the 
following 
conditions hold:
\begin{compactenum}
\item For each $1\leq j \leq n$ the 1-dimensional stochastic process 
$\{\vecseq{X}{i}{j}\}_{i=0}^{\infty}$ is adapted to 
$\{\genfilt_i\}_{i=0}^{\infty}$.
%\item For each $i\in \Nset_0$ and each $1\leq j \leq n$ it holds 
%$\E[|\vecseq{X}{i}{j}|]<\infty$.
\item For each $\omega \in \Omega$, $i\in \Nset_0$ and $1\leq j \leq n$ it holds 
$\vec{X}_i (\omega)[j]\geq {0}$ and $\vec{X}_i(\omega)[j]\neq \infty$.
\item For each $i\in \Nset_0$ there exists an almost-sure partition of the set $\{\stime>i\}$ into $n$ subsets $L^i_1,\dots,L^i_n$, all of them $\genfilt_i$-measurable, such that for each $1\leq j \leq n$:
\begin{compactitem}
	\item $\E[\vecseq{X}{i+1}{j}\mid 
	\genfilt_i]\leq\vecseq{X}{i}{j} - 
	\eps$ on $L_j$; and
	\item for all $1 \leq j' < j$ we have $\E[\vecseq{X}{i+1}{j'}\mid 
	\genfilt_i]=\vecseq{X}{i}{j'}$ on $L_{j}$.
\end{compactitem}
\end{compactenum}
\end{definition}

@@

For $\omega\in \Omega$ we define the level of $\omega$ at step $i$ to be the 
number $\levelrank{\omega}{i}$ equal to either $0$, if $\stime(\omega)\leq i$ or $\omega$ does not belong to $L^i_1 \cup \dots\cup L^i_n$, 
or otherwise to the $1\leq j \leq n$ such that $\omega\in L^i_j$. By the definition of 
$\eps$-LexRSM the value $\levelrank{\omega}{i}$ is well-defined for all 
$\omega$ and moreover, the random variable $\levelrank{}{i}$ is $\genfilt_i$-measurable. We denote by $\minlev(\omega)$ the smallest $0 \leq j \leq n$ such 
that $j$ is a level of $\omega$ at infinitely many steps. Note that $\omega \in 
\genRunSet_{\infty}$ implies $\minlev(\omega)\neq 0$, and $\probm(\genRunSet_{\infty})=\probm(\{\minlev \neq 0 \})$, as for each $i$ almost all $\omega$'s for which $\stime(\omega)>i$ belong to $L^i_1 \cup \dots\cup L^i_n$. We denote by $M_i$ 
the set of all $\omega$'s with $\minlev(\omega)=i$.
